Restructuring Data for Exascale Scientific Productivity

### 2016-03-08 papers/2016-Supercomputing/paper.tex
starting intro

As the exascale computing age emerges, data related issues are becoming
critical factors that determine how and where we do computing. Popular
approaches used by traditional I/O solution and storage libraries become
increasingly bottlenecked due to their assumptions about data movement,
re-organization, and storage. While, new technologies, such as ``burst
buffers'', can help address some of the short-term performance issues, it is
essential that we reexamine the underlying storage and I/O infrastructure to
effectively support requirements and challenges at exascale and beyond.

In this paper we present
- different data needs different techniques
- write overhead for the different refactoring/auditing approaches
- derived quantities in viz after precision based refactoring and differences
  from full fidelity data sets.
- linear auditing approaches for data refactoring and the effects on data
  sizes, io performance, and viz quality. Do every point, but split at 24 bits
and 40 bits.

@@

Be sure to address the following:\\
1. Why should Chang or Chen care from the apps side.\\
2. why should ken or hank care from the viz side\\
3. why would Garth care from the storage side\\

>>>>

  Big data and the associated changes in the computing platform are key
  components for the advances from exascale systems for scientific discovery
  with computing. Data will increasingly drive the next generation of scientific
  discoveries, but limitations in the rate of progress for data storage and
  movement technologies is likely to hamper this revolution. Storage performance
  and capacity bottlenecks have the potential to greatly limit how much data can
  be output from exascale simulations, how often this data can be output, and
  how much knowledge can be derived from the data.

  New techniques in storage and I/O hardware is one direction towards addressing
  these issues. In this paper we present an alternate method for addressing the
  data deluge, by utilizing domain specific knowledge about the data to manage
  the organization, content and precision of the output data. The tradeoff
  between system level quality of service and domain specific quality of
  information can be used to circumvent the storage bottlenecks without
  exploding the cost or energy usage of storage systems.

  In this paper we present a lossless data partitioning technique that
  intelligently utilizes hierarchical storage in next-generation leadership
  computing systems. We show that lower accuracy partitions can be used to
  service the need for most common use cases, and additional accuracy can be
  added for more demanding scenarios. The performance benefits are .....

  \begin{comment}
    As the exascale computing age emerges, data related issues are becoming
    critical factors that determine how and where we do computing. Popular
    approaches used by traditional I/O solution and storage libraries become
    increasingly bottlenecked due to their assumptions about data movement,
    re-organization, and storage. While, new technologies, such as ``burst
    buffers'', can help address some of the short-term performance issues, it is
    essential that we reexamine the underlying storage and I/O infrastructure to
    effectively support requirements and challenges at exascale and beyond.

    In this paper we present
    - different data needs different techniques
    - write overhead for the different refactoring/auditing approaches
    - derived quantities in viz after precision based refactoring and differences
    from full fidelity data sets.
    - linear auditing approaches for data refactoring and the effects on data
    sizes, io performance, and viz quality. Do every point, but split at 24 bits
    and 40 bits.
  \end{comment}

@@

\begin{comment}
  Be sure to address the following:\\
  1. Why should Chang or Chen care from the apps side.\\
  2. why should ken or hank care from the viz side\\
  3. why would Garth care from the storage side\\
\end{comment}

  Big data and the associated changes in the computing platform are key
  components for the advances from exascale systems for scientific discovery
  with computing. Data will increasingly drive the next generation of scientific
  discoveries, but limitations in the rate of progress for data storage and
  movement technologies is likely to hamper this revolution. Storage performance
  and capacity bottlenecks have the potential to greatly limit how much data can
  be output from exascale simulations, how often this data can be output, and
  how much knowledge can be derived from the data.

  New techniques in storage and I/O hardware is one direction towards addressing
  these issues. In this paper we present an alternate method for addressing the
  data deluge, by utilizing domain specific knowledge about the data to manage
  the organization, content and precision of the output data. The tradeoff
  between system level quality of service and domain specific quality of
  information can be used to circumvent the storage bottlenecks without
  exploding the cost or energy usage of storage systems.

  We have undertaken this effort as part of a larger rethinking of the role of
  storage in high performance computing applications, the SIRIUS project. In
  this paper we demonstrate the initial efforts towards creating precision
  aware I/O techniques for different types of large scale data sets. Our
  experiments show that selecting for output accuracy based on the intended
  usage for data can yeild substantial performance benefits. Moreover, we show
  that lossless data partitioning techniques can be utilize in lieu of lossy
  compression methods to utilize the hierarchical storage structure of
  next-generation leadership computing platforms. By partitioning data into
  buckets of varying precision, many usecases can be met with smaller, less
  accurate buckets. For other use cases where only the highest precision will
  suffice, the additional cost can be paid as and when needed.
